---
title: "VULCAN Analysis Combined"
author: "Andrew Holding"
date: "1/22/2018"
output:
  html_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: tango
    number_sections: yes
    toc: yes
    toc_depth: 3
always_allow_html: yes
---

```{r setup_master, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

#Introduction

To reduce the overall size of this project read count and network data is stored within Rdata files. These are avalible from https://github.com/andrewholding/VULCANSupplementary. The raw data is avalible from the gene expression omnibus if reprocessing is required. The following script will download all the required files and rebuild the network file. The Rmd files are designed to recreate the processed data from the raw data if the processed data is not found.  

```{bash download, eval=FALSE}

git clone https://github.com/andrewholding/VULCANSupplementary.git

#Rebuild network
cd ChIP-seq_ER/data/
cat brca-expmat.rda.part1 brca-expmat.rda.part2 > brca-expmat.rda

```


#Estrogen Receptor-alpha qPCR Analysis

```{r ER qRPCR, child = 'qPCR_ER/Manuscript_Figures.Rmd'}
```

#Estrogen Receptor-alpha ChIP-Seq and VULCAN Analysis

```{r ER ChIP-seq, child = 'ChIP-seq_ER/rmarkdown.Rmd'}
```

#Estrogen Receptor-alpha qPLEX-RIME Analysis

```{r ER qRIME, child = 'qPLEX-RIME_ER/02.Supplementary.Rmd'}
```

#GRHL2 ChIP-Seq Analysis

```{r GRHL2 Chip, child = 'ChIP-seq_GRHL2/Manuscript_Figures.Rmd'}
```

#GRHL2 qPLEX-RIME Analysis

```{r GRHL2 qRIME, child = 'qPLEX-RIME_GRHL2/rRunTMTanalysis.Rmd'}
```

#GRHL2 qPLEX-RIME Analysis
```{r GRHL2 eRNA, child = 'eRNA/plot_eRNA.Rmd'}
```

# Technical Session Info
The following code describes the R environment used to generate this document and will help making it fully reproducible should there be future updates in any of the packages.
```{r sessionInfo}
sessionInfo()
```

