---
title: "QRIME Scene Investigation - Report"
author: "Andrew Holding & Matt Eldridge"
date: "9 January 2017"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
header-includes:
- \usepackage{longtable}
- \usepackage{booktabs}
---

```{r configuration_report, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(cache = FALSE)
```
\newpage

# Differential binding

```{r differential_expression_paramters, include=FALSE}
significanceLevel <- 0.05
controlLogFoldChangeThreshold <- 1.0
```


```{r fit_linear_model, include = FALSE}

factors <- c("Group", "Run")

model <- as.formula(paste(c("~ 0", factors), collapse = " + "))

fittedLinearModels <- list()
fittedLinearModelsExcludingControls <- list()

for (analysis in names(combinedProteinData))
{
  cat(analysis, "\n", sep = "")
  data <- combinedProteinData[[analysis]]

  # log2 transform
  exprs(data) <- log2(exprs(data))

  design <- model.matrix(model, data = pData(data))

  colnames(design) <- colnames(design) %>%
    sub(pattern = "^Group", replacement = "") %>%
    gsub(pattern = " ", replacement = "_")

  fittedLinearModels[[analysis]] <- lmFit(data, design = design)

  samples <- pData(data) %>%
    rownames_to_column("Sample") %>%
    filter(Group != controlGroup) %>%
    select(Sample) %>%
    unlist(use.names = FALSE)

  data <- data[, sampleNames(data) %in% samples]

  design <- model.matrix(model, data = pData(data))

  colnames(design) <- colnames(design) %>%
    sub(pattern = "^Group", replacement = "") %>%
    gsub(pattern = " ", replacement = "_")

  fittedLinearModelsExcludingControls[[analysis]] <- lmFit(data, design = design)
}
```

```{r contrasts, include = FALSE}

contrasts <- contrastsTable$Contrast

reformatGroup <- function(x) gsub(x, pattern = " ", replacement = "_")

reformatContrasts <- function(x)
{
  x %>%
    gsub(pattern = " ", replacement = "_") %>%
    sub(pattern = "_-_", replacement = " - ") %>%
    sub(pattern = "_vs_", replacement = " - ")
}

reformattedContrasts <- reformatContrasts(contrasts)

fittedContrasts <- list()

for (analysis in names(combinedProteinData))
{
  cat(analysis, "\n", sep = "")
  fittedLinearModel <- fittedLinearModels[[analysis]]
  # fittedLinearModel <- fittedLinearModelsExcludingControls[[analysis]]
  contrastMatrix <- makeContrasts(contrasts = reformattedContrasts, levels = fittedLinearModel$design)
  fittedContrasts[[analysis]] <- contrasts.fit(fittedLinearModel, contrastMatrix)
  fittedContrasts[[analysis]] <- eBayes(fittedContrasts[[analysis]], trend = TRUE, robust = TRUE)
}
```

```{r create_contrast_results, include = FALSE}

createContrastResults <- function(
  msnset,
  fittedLinearModel,
  fittedLinearModelExcludingControls,
  fittedContrasts,
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
{
  reformattedContrast <- reformatContrasts(paste(contrastGroups, collapse = " - "))
  reformattedContrastGroups <- reformattedContrast %>% strsplit(" - ") %>% unlist
  reformatedControlGroup <- reformatGroup(controlGroup)

  results <- topTable(fittedContrasts, coef = reformattedContrast, number = Inf, sort.by = "none")

  fittedIntensities <- as.data.frame(fittedLinearModel$coefficients)
  # fittedIntensities <- as.data.frame(fittedLinearModelExcludingControls$coefficients)
  contrastFittedIntensities <- select(fittedIntensities, one_of(reformattedContrastGroups))
  results$logFC <- contrastFittedIntensities[, 1] - contrastFittedIntensities[, 2]

  fittedIntensities <- as.data.frame(fittedLinearModel$coefficients)
  contrastFittedIntensities <- select(fittedIntensities, one_of(reformattedContrastGroups))
  controlFittedIntensities <- fittedIntensities[, reformatedControlGroup]
  results$logFCcontrol = apply(contrastFittedIntensities - controlFittedIntensities, 1, max)

  samples <- pData(msnset) %>%
    mutate(Sample = rownames(.)) %>%
    filter(Group %in% c(contrastGroups, controlGroup)) %>%
    select(Sample) %>%
    unlist(use.names = FALSE)

  intensities <- exprs(msnset) %>%
    as.data.frame %>%
    select(one_of(samples))

  # add N column for the number of replicates, i.e. runs where intensity
  # is not NA for both groups
  n <- rep(0, nrow(intensities))
  for (run in runs)
  {
    samples <- pData(msnset) %>%
      mutate(Sample = rownames(.)) %>%
      filter(Group %in% contrastGroups) %>%
      filter(Run == run) %>%
      select(Sample) %>%
      unlist(use.names = FALSE)
    n <- n +
      intensities %>%
      select(one_of(samples)) %>%
      transmute(n = ifelse(complete.cases(.), 1, 0)) %>%
      unlist(use.names = FALSE)
  }

  results <- results %>%
    mutate(N = n) %>%
    bind_cols(fittedIntensities %>% select(one_of(rev(reformattedContrastGroups), controlGroup))) %>%
    bind_cols(intensities) %>%
    left_join(proteinInfo, by = "Protein")

  results <- bind_cols(
    results %>% select(Protein, Gene, Description, N),
    results %>% select(-Protein, -Gene, -Description, -N)
  )

  results 
}

```


```{r contrast_2, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

Here follows the analysis output for `r paste(contrastName, processingDescription, sep = ", ")`.

```{r results_2, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
```

## MA plot

```{r ma_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-bound proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

\newpage

## Volcano plot

```{r volcano_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-bound proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 50,
                  pointSize = 1.5))
```
\newpage

## P-value histogram

```{r pvalue_histogram_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```



## QQ plot

```{r qq_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```


\newpage

# Differential binding results tables

```{r contrast_results_thresholds, include = FALSE}
significanceLevel <- 0.05
logFoldChangeThreshold <- 1.0
controlLogFoldChangeThreshold <- 1.0
```

The following tables contain the top ranking differentially bound proteins for each comparison. Included are all proteins that reach a statistical signficance of `r significanceLevel` in terms of the adjusted p-value and those with an absolute log~2~ fold change of `r logFoldChangeThreshold` or above.

The IgG column gives the larger of the log~2~ fold changes for the two groups against the IgG control and an asterisk indicates specific binding where this log~2~ fold change is above a threshold of `r controlLogFoldChangeThreshold`. N is the number of replicates in which the protein was observed.

In all cases, peptide intensities were quantile normalized and measurements with missing values were removed prior to summarization at the protein level.

```{r contrast_results_table, include = FALSE}

contrastResultsTable <- function(results, contrastName,
                                 significanceLevel = 0.05,
                                 logFoldChangeThreshold = 1.0,
                                 controlLogFoldChangeThreshold = 1.0)
{
  resultsTable <- results %>%
    select(Protein, Gene, N, logFC, AveExpr, adj.P.Val, B, logFCcontrol)

  resultsTable$Specific <- ifelse(resultsTable$logFCcontrol >= controlLogFoldChangeThreshold, "*", "")

  resultsTable <- resultsTable %>%
    filter(adj.P.Val < 0.05 | abs(logFC) >= 1.0) %>%
    arrange(desc(abs(logFC)))

  colnames(resultsTable) <- c("Protein", "Gene", "N", "log2FC", "Avg Expr", "p-value", "B", "IgG", "")

  print(
    xtable(
      resultsTable,
      align = "lllrrrrrrl",
      display = c("d", "s", "s", "d", "f", "f", "g", "f", "f", "s"),
      caption = paste("Top ranking differentially bound proteins from the", contrastName, "comparison,  sorted by log2 fold change.")
    ),
    include.rownames = FALSE,
    tabular.environment = "longtable",
    floating = FALSE,
    add.to.row = addtorow
  )
}
```

\newpage

```{r results_table_contrast_1_config, include = FALSE}

contrastGroups <- strsplit2(contrast, " vs ")[1,]
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```





