---
title: "QRIME Scene Investigation"
author: "Andrew Holding & Matt Eldridge"
date: "9 January 2017"
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
header-includes:
- \usepackage{longtable}
- \usepackage{booktabs}
---

****************************

```{r configuration, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(cache = FALSE)
```

```{r load_libraries, include = FALSE, cache = FALSE}
library(tidyr)
library(dplyr)
library(tibble)
library(MSnbase)
library(RColorBrewer)
library(ggplot2)
library(kfigr)
library(gridExtra)
library(xtable)
library(limma)
```

```{r load_functions, include = FALSE, cache = FALSE}
source("Functions/plots.R")
source("Functions/normalization.R")
```

```{r xtable_options, include=FALSE, cache = FALSE}

options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)

addtorow          <- list()
addtorow$pos      <- list()
addtorow$pos[[1]] <- c(0)
addtorow$command  <- c(paste("\\hline \n",
                             "\\endhead \n",
                             "\\hline \n",
                             "{\\footnotesize Continued on next page} \n",
                             "\\endfoot \n",
                             "\\endlastfoot \n",sep=""))
```

```{r read_protein_information, include = FALSE}
proteinInfo <- read.delim("ReferenceData/protein_gene_description.txt", stringsAsFactors = FALSE)
```

```{r read_metadata, include = FALSE}
metadata <- read.delim("SampleData/metadata.txt", check.names = FALSE)
contrastsTable <- read.delim("SampleData/contrasts.txt", check.names = FALSE)
runs <- levels(metadata$Run)
groups <- levels(metadata$Group)
controlGroup <- "IgG"
```

```{r colours, include = FALSE}
groupColours <- c(brewer.pal(9, "Blues")[5:7], brewer.pal(9, "RdPu")[5:7], brewer.pal(9, "Greens")[6])
names(groupColours) <- groups
sampleColours <- groupColours[metadata$Group]
names(sampleColours) <- metadata$Sample
```

```{r read_intensities, include = FALSE}

peptideData <- list()

for (run in runs)
{
  cat(run, "\n", sep = "")

  data <- read.delim(paste("ProcessedData/", run, ".txt", sep = ""), stringsAsFactors = FALSE, check.names = FALSE)

  peptideCount <- data %>% select(Sequence, Modifications) %>% distinct %>% nrow
  cat("Peptides:", peptideCount, "\n", sep = "")
  if (peptideCount != nrow(data))
    stop("Error: multiple rows for at least one peptide/modification (is this PSM-level data?)")

  proteinCount <- data %>% select(Protein) %>% distinct %>% nrow
  cat("Proteins:", proteinCount, "\n", sep = "")

  sampleColumns <- metadata %>%
    filter(Run == run) %>%
    select(Group) %>%
    unlist(use.names = FALSE) %>%
    as.character

  data <- MSnSet(
    exprs = data %>% select(one_of(sampleColumns)) %>% as.matrix,
    fData = data %>% select(Sequence, Modifications, Protein),
    pData = metadata %>% filter(Run == run) %>% mutate(RowName = Group) %>% column_to_rownames("RowName")
  )

  peptideData[[run]] <- data
}
```

\newpage 

# SampleSheet


```{r samples_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
print(
  xtable(
    metadata %>% select(Run, Tag, Group) %>% spread(Run, Tag),
    caption = "Isobaric tags used for each sample (group) and run."
  ),
  include.rownames = FALSE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)

print(
  xtable(
    contrastsTable %>% select(Contrast),
    caption = "Configured Constrasts."
  ),
  include.rownames = FALSE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```


# Peptide intensity data

```{r filter_missing_intensities, include=FALSE}
peptideDataExcludingMissingValues <- list()
for (run in runs)
{
  cat(run, "\n", sep = "")
  data <- peptideData[[run]]
  data <- data[which(complete.cases(exprs(data))),]
  peptideDataExcludingMissingValues[[run]] <- data
}
```

```{r combine_peptide_intensities, include = FALSE}

# function for combining data for runs into single MSnSet object
combineRuns <- function(runData, runs)
{
  combinedData <- NULL

  for (run in runs)
  {
    data <- runData[[run]]

    sampleNames(data) <- pData(data) %>%
      transmute(Sample = paste(Run, Group, sep = ":")) %>%
      unlist %>%
      as.character

    featureNames(data) <- fData(data) %>%
      transmute(ID = paste(Sequence, Modifications, sep = "|")) %>%
      unlist %>%
      as.character()

    if (is.null(combinedData))
    {
      combinedData <- data
    } else
    {
      combinedData <- BiocGenerics::combine(combinedData, data)
    }
  }

  return(combinedData)
}

combinedPeptideData <- combineRuns(peptideData, runs)
combinedPeptideDataExcludingMissingValues <- combineRuns(peptideDataExcludingMissingValues, runs)
```

```{r peptide_protein_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

countProteins <- function(data) data %>% fData %>% select(Protein) %>% distinct %>% nrow

peptideData[["All"]] <- combinedPeptideData
peptideDataExcludingMissingValues[["All"]] <- combinedPeptideDataExcludingMissingValues

counts <- bind_rows(
  peptideData %>% sapply(nrow) %>% t %>% as.data.frame %>% mutate(label = "Peptides"),
  peptideDataExcludingMissingValues %>% sapply(nrow) %>% t %>% as.data.frame %>% mutate(label = "Peptides with no missing values"),
  peptideData %>% sapply(countProteins) %>% t %>% as.data.frame %>% mutate(label = "Proteins"),
  peptideDataExcludingMissingValues %>% sapply(countProteins) %>% t %>% as.data.frame %>% mutate(label = "Proteins with no missing values")
)
rownames(counts) <- counts$label
counts <- counts %>% select(-label)

peptideData[["All"]] <- NULL
peptideDataExcludingMissingValues[["All"]] <- NULL

print(
  xtable(
    counts,
    align = "lrrrr",
    caption = "Numbers of peptides and proteins observed in each run."
  ),
  include.rownames = TRUE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```

`r figr("intensity_plots", TRUE, type="Figure")` shows the distribution of intensities for each sample within each run.

```{r intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of intensities from each run.", fig.height=8, fig.width=5}

minIntensity <- -2
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideData[[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```

\newpage

# Normalization of intensity data

```{r imputation, include = FALSE}
peptideDataImputedMin <- list()
peptideDataImputedKNN <- list()
for (run in runs)
{
  peptideDataImputedMin[[run]] <- impute(peptideData[[run]], "min")
  peptideDataImputedKNN[[run]] <- impute(peptideData[[run]], "knn")
}
```


```{r normalization, include = FALSE}

peptideDataNormalized <- list()

peptideDataNormalized[["IncludeMissingValues_NoNormalization"]] <- peptideData
peptideDataNormalized[["ExcludeMissingValues_NoNormalization"]] <- peptideDataExcludingMissingValues
peptideDataNormalized[["ImputeMin_NoNormalization"]] <- peptideDataImputedMin
peptideDataNormalized[["ImputeKNN_NoNormalization"]] <- peptideDataImputedKNN

for (missingValueMethod in c("IncludeMissingValues", "ExcludeMissingValues", "ImputeMin", "ImputeKNN"))
{
  peptideDataForNormalization <- peptideDataNormalized[[paste(missingValueMethod, "NoNormalization", sep = "_")]]

  quantileNormalizationAnalysis <- paste(missingValueMethod, "QuantileNormalization", sep = "_")
  quantileNormalizationExcludingIgGAnalysis <-
    paste(missingValueMethod, "QuantileNormalizationExcludingIgG", sep = "_")
  scaleNormalizationAnalysis <- paste(missingValueMethod, "ScaleNormalization", sep = "_")
  scaleNormalizationExcludingIgGAnalysis <-
    paste(missingValueMethod, "ScaleNormalizationExcludingIgG", sep = "_")
  scaleNormalizationTopIgGAnalysis <- paste(missingValueMethod, "ScaleNormalizationTopIgG", sep = "_")

  peptideDataNormalized[[quantileNormalizationAnalysis]] <- list()
  peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationExcludingIgGAnalysis]] <- list()
  peptideDataNormalized[[scaleNormalizationTopIgGAnalysis]] <- list()

  numberOfPeptidesWithHighestIgG <- 10

  for (run in runs)
  {
    data <- peptideDataForNormalization[[run]]

    # quantile normalization
    peptideDataNormalized[[quantileNormalizationAnalysis]][[run]] <- normalize(data, method = "quantiles")
    
    # scale normalization
    peptideDataNormalized[[scaleNormalizationAnalysis]][[run]] <- normalizeMedianScaling(data)

    samples <- setdiff(sampleNames(data), controlGroup)

    # quantile normalization
    normalized <- normalize(data[,samples], method = "quantiles")
    peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]][[run]] <- data
    exprs(peptideDataNormalized[[quantileNormalizationExcludingIgGAnalysis]][[run]])[,samples] <-
      exprs(normalized)

    # scale normalization
    peptideDataNormalized[[scaleNormalizationExcludingIgGAnalysis]][[run]] <-
      normalizeMedianScaling(data, samples)

    # scale normalization using peptides with highest IgG intensities for computing sample medians
    controlIntensities <- data %>%
      exprs %>%
      as.data.frame %>%
      select(one_of(controlGroup)) %>%
      unlist %>%
      as.numeric
    controlIntensities[is.na(controlIntensities)] <- -Inf
    controlIntensityOrder <- rev(order(controlIntensities))
    peptideDataNormalized[[scaleNormalizationTopIgGAnalysis]][[run]] <-
      normalizeMedianScaling(data, rowsForCalculatingMedian = controlIntensityOrder[1:numberOfPeptidesWithHighestIgG])
  }
}
```

```{r normalized_intensity_plots, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Density plots of normalized intensities from each TMT run where QUANTILE normalization was applied to peptide intensities that include misssing values.", fig.height=6.5, fig.width=5}

analysis <- "IncludeMissingValues_QuantileNormalization"

minIntensity <- 0
maxIntensity <- 15

plots <- vector("list")

for (i in 1:length(runs))
{
  run <- runs[i]
  intensities <- peptideDataNormalized[[analysis]][[run]] %>% exprs %>% log2 %>% as.data.frame
  plots[[run]] <- intensityDistributionPlot(
    intensities,
    groups,
    groupColours,
    title = run,
    xlab=expression(log[2](intensity)),
    minIntensity = minIntensity,
    maxIntensity = maxIntensity,
    showLegend = (i == 1))
}

do.call("grid.arrange", c(plots, ncol = 1))
```


\newpage 

# Protein-level quantification


```{r summarize_protein_intensities, include=FALSE}

proteinData <- list()

for (analysis in names(peptideDataNormalized))
{
  peptideDataForSummarization <- peptideDataNormalized[[analysis]]

  proteinData[[analysis]] <- list()

  for (run in runs)
  {
    data <- peptideDataForSummarization[[run]]

    # exclude rows with missing values (will have no effect if imputation carried out)
    data <- data[which(complete.cases(exprs(data))),]

    data <- combineFeatures(data, groupBy = fData(data)$Protein, fun = "sum", na.rm = FALSE)
    fData(data) <- fData(data) %>% select(Protein)

    proteinData[[analysis]][[run]] <- data
  }
}
```

```{r combine_protein_intensities, include = FALSE}

combinedProteinData <- list()

for (analysis in names(proteinData))
{
  for (run in runs)
  {
    data <- proteinData[[analysis]][[run]]

    sampleNames(data) <- pData(data) %>%
      transmute(Sample = paste(Run, Group, sep = ":")) %>%
      unlist %>%
      as.character

    if (is.null(combinedProteinData[[analysis]]))
    {
      combinedProteinData[[analysis]] <- data
    } else
    {
      combinedProteinData[[analysis]] <- BiocGenerics::combine(combinedProteinData[[analysis]], data)
    }
  }
}
```

```{r replicate_counts, include=FALSE}

isRepresented <- function(x) any(!is.na(x))

getReplicateCounts <- function(data)
{
  metadata <- pData(data) %>% rownames_to_column("Sample")

  intensities <- data %>%
    exprs %>%
    as.data.frame %>%
    mutate(N = 0)

  for (run in runs)
  {
    samples <- metadata %>%
      filter(Run == run) %>%
      select(Sample) %>%
      unlist %>%
      as.character

    intensities <- intensities %>%
      mutate(N = N + apply(select(., one_of(samples)), 1, isRepresented))
  }

  replicateCounts <- intensities %>%
    select(N) %>%
    mutate(Count = 0) %>%
    group_by(N) %>%
    summarize_each(funs(length)) %>%
    ungroup %>%
    mutate(N = as.character(N))

  replicateCounts <- replicateCounts %>%
    bind_rows(data_frame(N = "total", Count = sum(replicateCounts$Count)))

  return(replicateCounts)
}

replicateCounts <- getReplicateCounts(combinedProteinData[["ImputeKNN_ScaleNormalization"]])

replicateCountTableData <- inner_join(
  getReplicateCounts(combinedProteinData[["ImputeKNN_ScaleNormalization"]]) %>%
    rename(`Including missing values` = Count),
  getReplicateCounts(combinedProteinData[["ExcludeMissingValues_ScaleNormalization"]]) %>%
    rename(`Excluding missing values` = Count),
  by = "N"
)

replicateCountTableData <- replicateCountTableData %>%
  as.data.frame %>%
  column_to_rownames(var = "N") %>%
  t %>%
  as.data.frame %>%
  rownames_to_column(var = "Runs/replicates")
```


```{r replicate_count_table, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
# colnames(replicateCounts) <- c("Runs/replicates", "Proteins (includes missing values)", "Proteins")
print(
  xtable(
    replicateCountTableData,
    # align = "llr",
    caption = "Numbers of proteins identified in differing numbers of runs and total number of proteins identified in all runs."
  ),
  include.rownames = FALSE,
  size = "\\setlength{\\tabcolsep}{12pt}"
)
```


# Principal Component Analysis


```{r pca_plot, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Principal Componenent Analysis for proteins sampled in all 3 runs. The PCA was based on protein-level data resulting from summation of quantile normalized peptide intensities in which missing values were imputed using KNN-based nearest neighbour averaging. The first two principal components are displayed.", fig.height=5}

data <- combinedProteinData[["ImputeKNN_QuantileNormalization"]]

metadata <- pData(data) %>% rownames_to_column("Sample")

scaledIntensities <- data %>%
  exprs %>%
  as.data.frame %>%
  filter(complete.cases(.))

for (run in runs)
{
  samples <- metadata %>%
    filter(Run == run) %>%
    select(Sample) %>%
    unlist %>%
    as.character

  summedIntensities <- scaledIntensities %>%
    select(one_of(samples)) %>%
    rowSums

  scaledIntensities <- scaledIntensities %>%
    mutate_each(funs(. / summedIntensities), one_of(samples))
}

print(pcaPlot(scaledIntensities, metadata$Sample, metadata$Group, groupColours, labels = metadata$Run, legend = TRUE))
```

\newpage

# Differential Binding

```{r differential_expression_paramters, include=FALSE}
significanceLevel <- 0.05
controlLogFoldChangeThreshold <- 1.0
```


```{r fit_linear_model, include = FALSE}

factors <- c("Group", "Run")

model <- as.formula(paste(c("~ 0", factors), collapse = " + "))

fittedLinearModels <- list()
fittedLinearModelsExcludingControls <- list()

for (analysis in names(combinedProteinData))
{
  cat(analysis, "\n", sep = "")
  data <- combinedProteinData[[analysis]]

  # log2 transform
  exprs(data) <- log2(exprs(data))

  design <- model.matrix(model, data = pData(data))

  colnames(design) <- colnames(design) %>%
    sub(pattern = "^Group", replacement = "") %>%
    gsub(pattern = " ", replacement = "_")

  fittedLinearModels[[analysis]] <- lmFit(data, design = design)

  samples <- pData(data) %>%
    rownames_to_column("Sample") %>%
    filter(Group != controlGroup) %>%
    select(Sample) %>%
    unlist(use.names = FALSE)

  data <- data[, sampleNames(data) %in% samples]

  design <- model.matrix(model, data = pData(data))

  colnames(design) <- colnames(design) %>%
    sub(pattern = "^Group", replacement = "") %>%
    gsub(pattern = " ", replacement = "_")

  fittedLinearModelsExcludingControls[[analysis]] <- lmFit(data, design = design)
}
```

```{r contrasts, include = FALSE}

contrasts <- contrastsTable$Contrast

reformatGroup <- function(x) gsub(x, pattern = " ", replacement = "_")

reformatContrasts <- function(x)
{
  x %>%
    gsub(pattern = " ", replacement = "_") %>%
    sub(pattern = "_-_", replacement = " - ") %>%
    sub(pattern = "_vs_", replacement = " - ")
}

reformattedContrasts <- reformatContrasts(contrasts)

fittedContrasts <- list()

for (analysis in names(combinedProteinData))
{
  cat(analysis, "\n", sep = "")
  fittedLinearModel <- fittedLinearModels[[analysis]]
  # fittedLinearModel <- fittedLinearModelsExcludingControls[[analysis]]
  contrastMatrix <- makeContrasts(contrasts = reformattedContrasts, levels = fittedLinearModel$design)
  fittedContrasts[[analysis]] <- contrasts.fit(fittedLinearModel, contrastMatrix)
  fittedContrasts[[analysis]] <- eBayes(fittedContrasts[[analysis]], trend = TRUE, robust = TRUE)
}
```

```{r create_contrast_results, include = FALSE}

createContrastResults <- function(
  msnset,
  fittedLinearModel,
  fittedLinearModelExcludingControls,
  fittedContrasts,
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
{
  reformattedContrast <- reformatContrasts(paste(contrastGroups, collapse = " - "))
  reformattedContrastGroups <- reformattedContrast %>% strsplit(" - ") %>% unlist
  reformatedControlGroup <- reformatGroup(controlGroup)

  results <- topTable(fittedContrasts, coef = reformattedContrast, number = Inf, sort.by = "none")

  fittedIntensities <- as.data.frame(fittedLinearModel$coefficients)
  # fittedIntensities <- as.data.frame(fittedLinearModelExcludingControls$coefficients)
  contrastFittedIntensities <- select(fittedIntensities, one_of(reformattedContrastGroups))
  results$logFC <- contrastFittedIntensities[, 1] - contrastFittedIntensities[, 2]

  fittedIntensities <- as.data.frame(fittedLinearModel$coefficients)
  contrastFittedIntensities <- select(fittedIntensities, one_of(reformattedContrastGroups))
  controlFittedIntensities <- fittedIntensities[, reformatedControlGroup]
  results$logFCcontrol = apply(contrastFittedIntensities - controlFittedIntensities, 1, max)

  samples <- pData(msnset) %>%
    mutate(Sample = rownames(.)) %>%
    filter(Group %in% c(contrastGroups, controlGroup)) %>%
    select(Sample) %>%
    unlist(use.names = FALSE)

  intensities <- exprs(msnset) %>%
    as.data.frame %>%
    select(one_of(samples))

  # add N column for the number of replicates, i.e. runs where intensity
  # is not NA for both groups
  n <- rep(0, nrow(intensities))
  for (run in runs)
  {
    samples <- pData(msnset) %>%
      mutate(Sample = rownames(.)) %>%
      filter(Group %in% contrastGroups) %>%
      filter(Run == run) %>%
      select(Sample) %>%
      unlist(use.names = FALSE)
    n <- n +
      intensities %>%
      select(one_of(samples)) %>%
      transmute(n = ifelse(complete.cases(.), 1, 0)) %>%
      unlist(use.names = FALSE)
  }

  results <- results %>%
    mutate(N = n) %>%
    bind_cols(fittedIntensities %>% select(one_of(rev(reformattedContrastGroups), controlGroup))) %>%
    bind_cols(intensities) %>%
    left_join(proteinInfo, by = "Protein")

  results <- bind_cols(
    results %>% select(Protein, Gene, Description, N),
    results %>% select(-Protein, -Gene, -Description, -N)
  )

  results 
}

```


```{r contrast_2, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_2, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)
```

```{r ma_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

```{r volcano_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 50,
                  pointSize = 1.5))
```

```{r pvalue_histogram_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```

```{r qq_plot_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```

\newpage

```{r contrast_3, include = FALSE}
contrastGroups <- c("ER 90min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_3, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

print(contrastGroups)

```

```{r ma_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

```{r volcano_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Volcano plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(volcanoPlot(results,
                  significanceLevel = significanceLevel,
                  minLogFoldChange = -2.0,
                  maxLogFoldChange = 2.75,
                  controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
                  minLogFoldChangeForLabelling = 1.5,
                  maxNumberLabelledProteins = 50,
                  pointSize = 1.5))
```

```{r pvalue_histogram_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("Histogram of p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=3.5, fig.width=4.5}
print(histogram(results, "P.Value", xlab = "p-value"))
```

```{r qq_plot_3, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("QQ plot of the adjusted p-values for the ", contrastName, " comparison (", processingDescription, ")", sep = ""), fig.height=5, fig.width=5}
q <- results %>%
  select(adj.P.Val) %>%
  filter(!is.na(adj.P.Val)) %>%
  arrange(adj.P.Val) %>%
  unlist %>%
  as.numeric
qqplot(-log10(ppoints(length(q))),
       -log10(q),
       xlab=expression(Expected~~-log[10](italic(q))),
       ylab=expression(Observed~~-log[10](italic(q))))
abline(0, 1, col = "red")
```


\newpage

```{r contrast_4, include = FALSE}
contrastGroups <- c("FOXA1 45min", "FOXA1 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_4, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

print(contrastGroups)

```

```{r ma_plot_4, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```

\newpage

```{r contrast_5, include = FALSE}
contrastGroups <- c("FOXA1 90min", "FOXA1 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r paste(contrastName, processingDescription, sep = ", ")`

```{r results_5, include = FALSE}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

print(contrastGroups)

```

```{r ma_plot_5, echo=FALSE, warning=FALSE, message=FALSE, fig.cap=paste("MA plot of the average intensity against log~2~ fold change for the ", contrastName, " comparison (", processingDescription, "). Top ranking differentially-expressed proteins with false discovery rate below 0.05 are highlighted in pink. Open circles indicate that the protein is non-specific from the IgG control comparison.", sep = ""), fig.height=6}
print(maPlot(results,
             significanceLevel = significanceLevel,
             controlLogFoldChangeThreshold = controlLogFoldChangeThreshold,
             minLogFoldChangeForLabelling = 1.5,
             maxNumberLabelledProteins = 50,
             pointSize = 1.5))
```


\newpage

# Differential binding results tables

```{r contrast_results_thresholds, include = FALSE}
significanceLevel <- 0.05
logFoldChangeThreshold <- 1.0
controlLogFoldChangeThreshold <- 1.0
```

The following tables contain the top ranking differentially bound proteins for each comparison. Included are all proteins that reach a statistical signficance of `r significanceLevel` in terms of the adjusted p-value and those with an absolute log~2~ fold change of `r logFoldChangeThreshold` or above.

The IgG column gives the larger of the log~2~ fold changes for the two groups against the IgG control and an asterisk indicates specific binding where this log~2~ fold change is above a threshold of `r controlLogFoldChangeThreshold`. N is the number of replicates in which the protein was observed.

In all cases, peptide intensities were quantile normalized and measurements with missing values were removed prior to summarization at the protein level.

```{r contrast_results_table, include = FALSE}

contrastResultsTable <- function(results, contrastName,
                                 significanceLevel = 0.05,
                                 logFoldChangeThreshold = 1.0,
                                 controlLogFoldChangeThreshold = 1.0)
{
  resultsTable <- results %>%
    select(Protein, Gene, N, logFC, AveExpr, adj.P.Val, B, logFCcontrol)

  resultsTable$Specific <- ifelse(resultsTable$logFCcontrol >= controlLogFoldChangeThreshold, "*", "")

  resultsTable <- resultsTable %>%
    filter(adj.P.Val < 0.05 | abs(logFC) >= 1.0) %>%
    arrange(desc(abs(logFC)))

  colnames(resultsTable) <- c("Protein", "Gene", "N", "log2FC", "Avg Expr", "p-value", "B", "IgG", "")

  print(
    xtable(
      resultsTable,
      align = "lllrrrrrrl",
      display = c("d", "s", "s", "d", "f", "f", "g", "f", "f", "s"),
      caption = paste("Top ranking differentially bound proteins from the", contrastName, "comparison,  sorted by log2 fold change.")
    ),
    include.rownames = FALSE,
    tabular.environment = "longtable",
    floating = FALSE,
    add.to.row = addtorow
  )
}
```

\newpage

```{r results_table_contrast_1_config, include = FALSE}
contrastGroups <- c("ER 45min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_1, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_2_config, include = FALSE}
contrastGroups <- c("ER 90min", "ER 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_2, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_3_config, include = FALSE}
contrastGroups <- c("ER 90min", "ER 45min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_3, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_4_config, include = FALSE}
contrastGroups <- c("FOXA1 45min", "FOXA1 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_4, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_5_config, include = FALSE}
contrastGroups <- c("FOXA1 90min", "FOXA1 0min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_5, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/",paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```

\newpage

```{r results_table_contrast_6_config, include = FALSE}
contrastGroups <- c("FOXA1 90min", "FOXA1 45min")
contrastName <- paste(contrastGroups, collapse = " vs ")
processingDescription <- "excluding peptides with missing intensities, quantile normalization"
analysis <- "ExcludeMissingValues_QuantileNormalization"
```

## `r contrastName`

```{r results_table_contrast_6, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}

data <- combinedProteinData[[analysis]]
exprs(data) <- log2(exprs(data))

results <- createContrastResults(
  data,
  fittedLinearModels[[analysis]],
  fittedLinearModelsExcludingControls[[analysis]],
  fittedContrasts[[analysis]],
  runs,
  contrastGroups,
  controlGroup,
  proteinInfo)

contrastResultsTable(results, contrastName,
                     significanceLevel = 0.05,
                     logFoldChangeThreshold = 1.0,
                     controlLogFoldChangeThreshold = 1.0)

resultsFile <- paste("Results/", paste(gsub(" ", "_", contrastGroups), collapse = "_vs_"), analysis, ".txt", sep = "")
write.table(results, resultsFile, sep = "\t", row.names = FALSE, quote = FALSE)
```



```{r save, eval = TRUE, include = FALSE, cache = FALSE}
save.image(file = "Results/QRIME.RData")
```

